---
title: "Coursera week 2 work"
author: "yu yang liu"
date: "2017/2/5"
output: html_document
---
```{r}
knitr::opts_chunk$set(cache = T,echo = T)
```


##Loading and preprocessing the data
* unzip file
```{r} 
if(!file.exists("activity.csv")){
  unzip("repdata_data_activity.zip")
}
```
* load packages
```{r}
library(ggplot2)
library(dplyr)
library(data.table)
```
* load data
```{r}
df<-read.csv("activity.csv")
```
* view into data
```{r}
dim(df)
str(df)
head(df)
tail(df)
```


##What is mean total number of steps taken per day?

```{r,fig.width=6}
q2<-df%>%
  group_by(date)%>%
  summarize(total=sum(na.omit(steps)),mean=mean(na.omit(steps)),median=median(na.omit(steps)))

head(q2)    #view into data

ggplot(q2,aes(date,total))+
  geom_histogram(stat = "identity")+
  labs(title="Histogram of the total number of steps taken each day",y="total steps")+
  theme(axis.text.x = element_text(angle = 90))

ggplot(q2,aes(date,mean))+
  geom_histogram(stat="identity")+
  labs(title="Histogram of the mean of steps taken each day",y="mean steps")+
  theme(axis.text.x = element_text(angle = 90))

ggplot(q2,aes(date,median))+
  geom_histogram(stat="identity")+
  labs(title="Histogram of the median of steps taken each day",y="median steps")+
  theme(axis.text.x = element_text(angle = 90))
```

##What is the average daily activity pattern?

```{r q5}
q5<-df%>%
  group_by(interval)%>%
  summarize(mean=mean(na.omit(steps)))
q51<-q5$interval[q5$mean==max(q5$mean)]
q52<-max(q5$mean)

ggplot(q5,aes(interval,mean))+
  geom_line(stat="identity")+
  geom_vline(xintercept =q51,lty=6,color="red3",lwd=1.2,alpha=.5)+
  annotate("segment",x=q51+200,xend =q51+25,y=q52,yend=q52,arrow=arrow(),size=.75)+
  annotate("text",x=q51+620,y=q52,label=paste("steps max:",round(q52,2)),size=6)+
  scale_x_continuous(breaks = c(0,500,q51,1000,1500,2000))+
  labs(y="mean steps",title="Average daily activity pattern")
```

we can see when the time is at 8:30, it have the maximum steps taken.  
About 206.17 steps taken peer days

##Imputing missing values

First we recap what the data look like
```{r q6}
head(df)
tail(df)
```


* how many NA in data
```{r}
q61<-c(length(which(is.na(df[,1]))),round(length(which(is.na(df[,1])))/nrow(df),2))
q62<-length(which(is.na(df[,2])))
q63<-length(which(is.na(df[,3])))
q65<-q61+q62+q63
q66<-data.frame(steps=q61,date=q62,interval=q63,total=q65)
rownames(q66)<-c("quantity", "percent (%)")
q66
```
there have 2304 NA in "steps", about 0.13% of total data.  
other variables have no missing data

* deal with missing values
```{r}
q64<-df%>%
  group_by(interval)%>%
  summarize(max=mean(na.omit(steps)))
head(q64)
```

I want to fill missing values with the mean for that 5-minute interval
```{r}
for(i in 1:nrow(df)){
    if(is.na(df[i,1])){
      a<-which(q64[,1]==df[i,3])
    df[i,1]<-q64[a,2]
  }
}
head(df)
```
so the new data would like above  
  
Histogram of the total number/mean/median of steps taken each day after missing values are imputed would look like this
```{r q7}

q7<-df%>%
  group_by(date)%>%
  summarize(total=sum(steps),mean=mean(steps),median=median(steps))

ggplot(q7,aes(date,total))+
  geom_histogram(stat = "identity")+
  labs(title="Histogram of the total number of steps taken each day",y="total steps")+
  theme(axis.text.x = element_text(angle = 90))

ggplot(q7,aes(date,mean))+
  geom_histogram(stat="identity")+
  labs(title="Histogram of the mean steps taken each day",y="mean steps")+
  theme(axis.text.x = element_text(angle = 90))

ggplot(q7,aes(date,median))+
  geom_histogram(stat="identity")+
  labs(title="Histogram of the median steps taken each day",y="median steps")+
  theme(axis.text.x = element_text(angle = 90))

```



##Are there differences in activity patterns between weekdays and weekends?
* indicating whether the day data give is a weekday or weekend day
```{r q8}
q8<-df%>%
  mutate(week=weekdays(as.POSIXct(df$date)))

q8<-q8%>%
  mutate(week2=ifelse(week%in%c("星期一","星期二","星期三","星期四","星期五"),"weekdays","weekends"))

```

the levels of weekdays/weekends variable
```{r}
levels(as.factor(q8$week2))
```

* panel plot containing a time series plot of the 5-minute interval and the average number of steps taken

```{r}
q81<-q8%>%
  group_by(week2,interval)%>%
  summarize(mean=mean(steps))
ggplot(q81,aes(interval,mean))+
  geom_line(stat="identity")+
  labs(title="time series plot for average steps by weekdays/weekends",y="mean of steps")+
  facet_grid(week2~.)
```

